---
title: "Gym Profit"
author: "Cameron Cheatham"
date: "2025-07-20"
output:
  pdf_document: default
  html_document: default
---
## Gym Profit Report

This report analyzes the gym's profit over six months based on membership growth and pricing strategy.

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load necessary library
library(tidyverse)

# Define months
months <- c("July", "August", "September", "October", "November", "December")

# Membership numbers
standard_members <- c(34, 39, 45, 90, 180, 360)
premium_members <- c(33, 67, 133, 266, 532, 1064)

# Pricing and costs
standard_price <- 20
premium_price <- 60
variable_cost_per_member <- 6
fixed_costs <- 6586

# Calculate totals and profit
total_members <- standard_members + premium_members
total_revenue <- (standard_members * standard_price) + (premium_members * premium_price)
variable_costs <- total_members * variable_cost_per_member
total_costs <- variable_costs + fixed_costs
profit <- total_revenue - total_costs

# Create data frame with factors to keep order
gym_data <- data.frame(
  Month = factor(months, levels = months),
  Standard_Members = standard_members,
  Premium_Members = premium_members,
  Total_Members = total_members,
  Total_Revenue = total_revenue,
  Variable_Costs = variable_costs,
  Fixed_Costs = fixed_costs,
  Total_Costs = total_costs,
  Profit = profit
)

# Print the data frame
print(gym_data)
```

## Profit Over Time
The chart below visualizes the gym's profit over the six months. The red dashed line shows the break-even point at zero profit.

```{r profit-chart, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)

ggplot(gym_data, aes(x = Month, y = Profit, group = 1)) +
  geom_line(color = "blue", size = 1.2) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Gym Profit Over Time",
       x = "Month",
       y = "Profit ($)") +
  theme_minimal()
```

## Members vs Profit
This chart shows the relationship between total members and profit over time.

```{r gym-chart, echo=FALSE, warning=FALSE, message=FALSE} 
ggplot(gym_data, aes(x = Month, y = Profit, group = 1)) +
  geom_line(color = "darkgreen", size = 1.2) +              # Profit line
  geom_point(size = 3) +                               # Profit points
  geom_smooth(method = "lm", se = FALSE, color = "blue", linetype = "dashed") +  # Trend line
  geom_text(aes(label = Total_Members),                # Add total members as text labels
            vjust = -1.2, size = 3.2, color = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Gym Profit Over Time with Total Members",
       y = "Profit ($)",
       x = "Month") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )
```

## Key Insights
- The gym starts generating profit in **September**.
- Membership must grow by **20% each month** to achieve this.
- By **December**, profit reaches over **$55,000** if the trend continues.
- A mix of **Standard** and **Premium** members improves revenue.
