---
title: "Gym Expansion Feasibility Study – Phase 2"
author: "Cameron Cheatham"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load necessary packages
library(tidyverse)
library(ggplot2)
library(scales)
library(formattable)
library(tidyr)
```

#key financial variables:
- expansion_cost: Total estimated cost to expand the gym.
- expenses: Current annual expenses.
- revenue: Current annual revenue.
- growth_rate: Expected annual growth rate for profit.
- net_profit: Profit calculated as revenue minus expenses.


```{r echo=FALSE}

expansion_cost <- 150000
expenses <- 96900
revenue <- 140000
growth_rate <- 0.25
net_profit <- revenue - expenses
```

# This section calculates:
- The required annual profit to cover the expansion cost over 3 years.
- The compound annual growth rate (CAGR) needed to meet expansion costs based on current profits.
- The projected revenue after 3 years, assuming growth at CAGR.

```{r echo=FALSE, results='asis'}
required_annual_profit <- expansion_cost / 3
cagr <- (expansion_cost / net_profit)^(1/3) - 1
projected_revenue <- revenue * (1 + cagr)^3

cat("**Required Annual Profit:**", dollar(required_annual_profit), "\n")
cat("**Compound Annual Growth Rate (CAGR):**", round(cagr * 100, 2), "%\n")
cat("**Projected Revenue in 3 Years (at CAGR):**", dollar(round(projected_revenue, 2)), "\n")
```

## 3-Year Financial Forecast
This forecast projects revenue, expenses, and net profit over the next 3 years based on:
- Price increase of 10% in Year 1.
- Membership growth of 15% in Year 2.
- Expense growth and cost savings in Years 2 and 3.

```{r echo=FALSE}
forecast_years <- c("Y1", "Y2", "Y3")

forecast_revenue <- c(
  revenue * 1.10,                    # Y1: +10% price
  revenue * 1.10 * 1.15,             # Y2: +15% members
  revenue * 1.10 * 1.15              # Y3: same as Y2
)

forecast_expenses <- c(
  expenses,                          # Y1: same expenses
  expenses * 1.10,                   # Y2: +10% growth costs
  expenses * 1.10 * 0.95             # Y3: -5% cost savings
)

forecast_net_profit <- forecast_revenue - forecast_expenses

forecast_notes <- c(
  "Raise prices by 10%",
  "Add 15% more members",
  "Cut costs by 5%"
)

forecast_df <- data.frame(
  Year = forecast_years,
  Revenue = round(forecast_revenue, 2),
  Expenses = round(forecast_expenses, 2),
  Net_Profit = round(forecast_net_profit, 2),
  Notes = forecast_notes
)

knitr::kable(forecast_df, caption = "3-Year Financial Forecast")
```
```{r echo=FALSE}
ggplot(forecast_df, aes(x = Year, y = Net_Profit, group = 1)) +
  geom_line(color = "steelblue", linewidth = 1.2) +
  geom_point(color = "darkred", size = 3) +
  geom_hline(yintercept = expansion_cost, linetype = "dashed", color = "gray40") +
  labs(
    title = "3-Year Forecast: Net Profit vs. Expansion Cost",
    y = "Net Profit ($)",
    x = "Year"
  ) +
  theme_minimal()
```

### 3-Year Financial Forecast
This 3-year forecast projects revenue, expenses, and net profit based on realistic assumptions such as price increases, membership growth, and cost savings.
```{r echo=FALSE, fig.align='center', fig.width=6, fig.height=3}
forecast_long <- forecast_df %>%
  pivot_longer(cols = c("Revenue", "Expenses", "Net_Profit"),
               names_to = "Metric", values_to = "Amount")

ggplot(forecast_long, aes(x = Year, y = Amount, fill = Metric)) +
  geom_col(position = "dodge", width = 0.5) +
  labs(
    title = "3-Year Financial Forecast: Revenue, Expenses, and Net Profit",
    y = "Amount ($)",
    x = "Year",
    fill = "Metric"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7.5))
```

**However, the 3-year net profit projections indicate it is unlikely to fully cover the gym expansion cost within this timeframe.** This suggests that a longer period is needed to recoup the investment.
```{r echo=FALSE}
growth_rate <- 0.25
years <- 0:6

profit_projection <- net_profit * (1 + growth_rate)^years

profit_df <- data.frame(
  Year = years,
  Projected_Profit = round(profit_projection, 2)
)

knitr::kable(profit_df, caption = "7-Year Profit Projection at 25% Growth")
```

# 7-Year Profit Projection at 25% Growth
Given the 3-year forecast is insufficient to cover expansion costs, we extend the projection to 7 years assuming a steady 25% annual growth in profits.

This longer-term projection helps evaluate if the expansion becomes financially feasible over a more realistic timeline.
```{r echo=FALSE}
ggplot(profit_df, aes(x = Year, y = Projected_Profit)) +
  geom_line(color = "darkgreen", linewidth = 1.2) +
  geom_point(aes(color = Year < 6), size = 3) +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "darkgreen")) +
  labs(
    title = "Projected Profit Growth at 25% Annual Rate",
    x = "Year",
    y = "Projected Profit ($)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

# Summary and Conclusion

In this feasibility study, we evaluated the financial outlook for expanding the gym.

- The **3-year forecast** showed that while revenue and profit grow, it is unlikely the expansion cost will be fully covered within this short timeframe.
- Extending the projection to **7 years**, assuming a consistent 25% annual growth rate, indicates the investment becomes achievable with sustained growth.
- This suggests the expansion is a **long-term commitment**, requiring strategic planning to maintain growth and manage costs.
- Moving forward, focusing on **increasing membership, optimizing pricing, and controlling expenses** will be critical to achieving these financial goals.

Overall, while immediate payback within 3 years is not feasible, the 7-year projection provides a promising outlook for successful expansion.


